Should I Airbnb My Property?

### 2017-03-03 abb_Writeup.Rmd
adding host block images

   load(paste0(data.path, '/analyzed/analyzed_workspace.RData'))

@@

* An investor purchases a property in Melbourne, Australia on September 1, 2015.
* They are looking to maximize the net revenue over the 12-month period from September 1, 2015 to August 31, 2016.
* The choices are 1) traditional year long rental (long term), or: 2) short-term, nightly Airbnb-type rental.
* Only direct monetary revenues and costs are considered, others such as liquidity and inconvenience are not considered.
* As most long-term rental use a leasing agency, we will assume use of an agency.

@@

<!-- ### Summary of Data Cleaning -->

<!-- We have applied nine separate filters to our two datasets.  In the name of transparency and to highlight the bounds of our research questions it is useful to review this filters (Table 2). There are more than 20,000 properties that are or have been listed on the Airbnb portal in Melbourne during the 2014 to 2016 period.  Filtering those not listed during the study period removes 1/3 of all properties.  Further removing by property type and listing type (not entire home) elimates another 50% of the remaining properties.  From these 6,600 properties structural characteristics, locational issues and outlying nightly rates remove another 800 from consideration.  Finally, 1,400 of these have never been booked and have \$0 revenue, requiring removal.  From over 20,000, we have 4,446 Airbnb properties that meet the requirements of our research question. -->

<!-- For the long term data, over half of the observations fall outside of our considered time frame, reducing the sample size down to around 130,000 from 290,000.  Further consideration regarding property types, structural characteristics, location, etc. remove another 10,000 observations, leaving 122,051 in the final dataset. -->

@@

rownames(clean.table) <- NULL
clean.table$field <- c('Initial Count', 'Time', 'Property Type',
                       'Listing Type', 'Bedrooms', 'Baths', 'Bed/Bath Combo',
                       'Rates & Rents', 'Location', 'Revenue')
clean.table$strper <- 100*round(clean.table$strper, 2)
clean.table$ltrper <- 100*round(clean.table$ltrper, 2)
names(clean.table) <- c('Filter', 'Short Term', 'Cut', '% Cut',
                        'Long Term', 'Cut', '% Cut')
options(xtable.comment = FALSE)
print(xtable(clean.table, caption='Data Cleaning Summary\n',
       digits=0),
       include.rownames=FALSE,
       caption.placement='top')

@@

<!-- ## Comparison -->

<!-- #### Observed Revenue -->

<!-- Our final data filtering considers the observed gross revenue of the properties. For the long-term rentals the observed gross revenue is the weekly rate multiplied by 52 minus the number of weeks the property was on the market. In a number of extreme cases, time on market was extremely long and the total observed revenues very low. To remove these outlying observations, we filter out any properties with less than $10,000 in revenue. -->

<!-- For the short-term data, we first calculate the total number of bookings, the occupancy rate, blocked rate and proportion of the year over which the property was listed.  Using this data we sum up the total revenue from bookings. For properties that were not listed on the site for the entire year, we impute their annual revenue by dividing the actual revenue by the proportion of the year in which they were listed.^[Properties listed in summer, but not winter may be over imputed and vice versa, however, we expect the biases to cancel out over the entire sample.] Properties with no reserved bookings, and, therefore, no revenue, are removed. -->

@@

<!-- ```{r revdensity.plot, message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE, echo=FALSE, fig.width=6,fig.height=3.5, fig.cap='Comparison of Revenue Distributions'} -->

<!--   strr <- str.data[,c('bedrooms', 'gross.revenue')] -->
<!--   strr$tenure <- 'short-term' -->
<!--   ltrr <- ltr.data[,c('bedrooms', 'gross.revenue')] -->
<!--   ltrr$tenure <- 'long-term' -->
<!--   allr <- rbind(strr, ltrr) -->

<!--   ggplot(allr, aes(x=gross.revenue, fill=tenure, color=tenure)) + -->
<!--     geom_density(alpha=.5) + -->
<!--     scale_fill_manual(values=c(ltr.col, abb.col)) + -->
<!--     scale_color_manual(values=c(ltr.col, abb.col)) + -->
<!--     ggtitle('Annual Revenues by Tenure Type') + -->
<!--     xlab('Annual Revenue') + -->
<!--     scale_x_continuous(breaks=c(seq(0,60000, by=20000)), -->
<!--                        labels=c('$0', '$20,000', '$40,000', '$60,000'))+ -->
<!--     theme(legend.position='bottom', -->
<!--           legend.title = element_blank(), -->
<!--           plot.title = element_text(hjust = 0.5), -->
<!--           axis.text.y = element_blank(), -->
<!--           axis.ticks.y = element_blank(), -->
<!--           axis.title.y=element_blank()) + -->
<!--     coord_cartesian(xlim=c(0, 70000)) -->

@@

<!-- <!-- Although it gathers considerable interest in both economic and public policy circles, analyses of Airbnb market and their comparison to more traditional long-term leasing strategies remains relatively unaddressed.  This work attempts to fill this void.  --> -->

>>>>

   load(paste0(data.path, '/analyzed/analyzed_workspace.RData'))
a <- 1

@@

* An investor purchases a property in Melbourne, Australia on September 1, 2015.
* They are looking to maximize the net revenue over the 12-month period from September 1, 2015 to August 31, 2016.
* The choices are 1) traditional year long rental (long term), or: 2) short-term, nightly Airbnb-type rental.
* Only direct monetary revenues and costs are considered, others such as liquidity and inconvenience are not considered.
* As most long-term rentals use a leasing agency, we will assume use of an agency.

@@

### Summary of Data Cleaning

We have applied ten separate filters to our two datasets. In the name of transparency and to highlight the bounds of our research questions it is useful to review these filters (Table 2). There are more than 20,000 properties that are or have been listed on the Airbnb portal in Melbourne during the 2014 to 2016 period.  Filtering those not listed during the study period removes nearly 1/3 of all properties.  Further removing by property type and listing type (not entire home) elimates another 50% of the remaining properties.  From these 7,200 properties structural characteristics remove another 480 from consideration.  Over 1,000 of the remaining properties were never been booked and/or had outlying nightly rates and were removed.  Finally, 14 properties had either bad location data or missing information host policy. From over 20,000, we have 5,788 Airbnb properties meet the requirements of our research question.

For the long term data, over half of the observations fall outside of our considered time frame, reducing the sample size down to around 130,000 from 290,000.  Further consideration regarding property types, structural characteristics, location, etc. remove another 12,000 observations, leaving 120,092 in the final dataset.

@@

rownames(clean.table) <- NULL
clean.table$field <- c('Initial Count', 'Time', 'Property Type',
                       'Listing Type', 'Bedrooms', 'Baths', 'Bed/Bath Combo',
                       'Rates & Rents', 'Revenue', 'Location', 'Host Policy')
clean.table$strper <- 100*round(clean.table$strper, 2)
clean.table$ltrper <- 100*round(clean.table$ltrper, 2)
names(clean.table) <- c('Filter', 'Short Term', 'Cut', '% Cut',
                        'Long Term', 'Cut', '% Cut')
options(xtable.comment = FALSE)
print(xtable(clean.table, caption='Data Cleaning Summary\n',
       digits=0),
       include.rownames=FALSE,
       caption.placement='top')

@@

## Revenues, More Explanation

#### Revenue Calculations

Calculations of revenue for the observed and hypothetical long-term rentals are straightforward, weekly rent -- real or imputed -- times 52 minus search cost time and agency fees.  For the short-term rental it is somewhat more complicated.  First, we must deal with those properties that were not listed for the entire 12 month period.  Our data cleaning eliminated all properties that were listed for less than six months, however, it would be incorrect to compare the revenue of a property listed on Airbnb for seven months with the hypothetical annual revenue from a long term rental.

Second, many of the short term properties have a significant number of blocked days, days in which the property was not offered on Airbnb.  Here too, there is a problem with comparing observed revenues from a property with a 40\% block rate with a hypothetical year long revenue stream, especially if the property was leased out on other portal during those blocked days.  And, even if we knew that the blocked days were simply due to the host wanting to use the property themselves during that time, this deduction in revenue should not necessarily apply to the hypothetical investor in our examined scenario.  In short, in either case, the actual observed short-term revenue would be biased on the low end compared to what a revenue maximizing owner would be likely able to accomplish.

As a result, we have calculated three separate revenues values for the short-term properties:

1. \textbf{Actual ($REV_a$)}: Observed bookings multiplied by median nightly rate
2. \textbf{Extrapolated ($REV_e$)}: Observed bookings extrapolated to a full year^[Properties listed in summer, but not winter may be over imputed and vice versa, however, we expect the biases to cancel out over the entire sample.] based on observed occupancy rate multiplied by median nightly rate
3. \textbf{Potential ($REV_p$)}: Observed occupancy rate of non-blocked days multiplied by 366 multiplied by median nightly rate

@@

```{r revdensity.plot, message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE, echo=FALSE, fig.width=6,fig.height=3.5, fig.cap='Comparison of Revenue Distributions'}

  stre <- str.data[,c('bedrooms', 'str.ext.revenue')]
  stre$tenure <- 'short-term extrapolated   '
  names(stre)[2] <- 'revenue'
  strp <- str.data[,c('bedrooms', 'str.pot.revenue')]
  strp$tenure <- 'short-term potential   '
  names(strp)[2] <- 'revenue'
  ltrr <- ltr.data[,c('bedrooms', 'gross.revenue')]
  ltrr$tenure <- 'long-term   '
  names(ltrr)[2] <- 'revenue'
  allr <- rbind(stre, strp, ltrr)

  ggplot(allr, aes(x=revenue, fill=tenure, color=tenure)) +
    geom_density(alpha=.5) +
    scale_fill_manual(values=c(ltr.col, abb.col[2], str.col)) +
    scale_color_manual(values=c(ltr.col, abb.col[2], str.col)) +
    ggtitle('Annual Revenues by Tenure Type') +
    xlab('Annual Revenue') +
    scale_x_continuous(breaks=c(seq(0,60000, by=20000)),
                       labels=c('$0', '$20,000', '$40,000', '$60,000'))+
    theme(legend.position='bottom',
          legend.title = element_blank(),
          plot.title = element_text(hjust = 0.5),
          axis.text.y = element_blank(),
          axis.ticks.y = element_blank(),
          axis.title.y=element_blank()) +
    coord_cartesian(xlim=c(0, 70000))

```
```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=FALSE}

  #host.block.freq <- 
  ggplot(str.data, aes(x=block.rate)) + 
  geom_density(fill=str.col, color=str.col) +
  ylab('Frequency\n') + 
  xlab('\n% of Time Property is Blocked') +
  scale_x_continuous(breaks=seq(0,1,.2),
                     labels=c('0%', '20%', '40%', '60%', '80%', '100%')) +
    ggtitle('Host Block Rate Frequency\n') +
    theme(plot.title = element_text(hjust = 0.5)) +
    theme(axis.text.y=element_blank(),
          axis.ticks.y=element_blank())
```

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=FALSE}

  #block.rate.period <- 
ggplot(str.data, aes(x=block.rate, y=nbr.block*extr.par)) + 
  geom_point(fill=str.col, color=str.col, size=.2) +
  ylab('Nbr of Blocked Periods\n') + 
  xlab('\n% of Time Property is Blocked') +
  scale_x_continuous(breaks=seq(0, 1, .2),
                     labels=c('0%', '20%', '40%', '60%', '80%', '100%')) +
    ggtitle('Block Rate vs. Nbr. of Blocked Periods\n') +
    theme(plot.title = element_text(hjust = 0.5))

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=FALSE}

str.data$blockpertime <- round(str.data$nbr.block * str.data$extr.par, 1)

str.data$host.type <- 'Unknown'
str.data$host.type[str.data$blockpertime > 12] <- 'Multi-Platform User'
str.data$host.type[str.data$block.rate <= .25] <- 'Profit Seeker'
str.data$host.type[str.data$block.rate >= .75] <- 'Opportunistic Sharer'
str.data$host.type <- factor(str.data$host.type,
                             levels=c('Profit Seeker', 'Opportunistic Sharer', 
                                      'Multi-Platform User', 'Unknown'))

#host.type <-

ggplot(str.data, 
       aes(x=block.rate, y=blockpertime, color=host.type)) + 
  geom_point(alpha=.5, size=.8, shape=16) +
  ylab('Number of Blocked Periods\n') + 
  xlab('\n% of Time Property is Blocked') +
      ggtitle('Block Rate vs. Nbr. of Blocked Periods\n') +

    scale_x_continuous(breaks=seq(0, 1, .2),
                     labels=c('0%', '20%', '40%', '60%', '80%', '100%')) +

  scale_color_manual(values=c(abb.col[2], abb.col[4], abb.col[7], abb.col[9])) + 
  theme(legend.position='bottom',
          legend.title = element_blank(),
        plot.title = element_text(hjust = 0.5))+
    guides(colour = guide_legend(override.aes = list(size=3,
                                                     alpha=1)))

## Comparison

@@

<!-- <!-- Although it gathers considerable interest in both economic and public policy circles, analyses of Airbnb market and their comparison to more traditional long-term leasing strategies remains relatively unaddressed.  This work attempts to fill this void.  -->
